Quasi-Newton Methods Comparision using Ackley, Beale, Booth, Matyas, Rastrigin and RosenBrock Functions

### 2022-12-30 paper/new/article.tex
Article Done

\begin{itemize}
	\item \textit{Convergence: } the convergence of an algorithm is classified as \textit{good}, \textit{poor} and \textit{divergence}.
	\subitem \textit{Good convergence: } the final solution is really near the optimal solution within the max iterations;
	\subitem \textit{Poor convergence: } the final solution is far from the optimal solution for any reason;
	\subitem \textit{Divergence: }  the final solution is out of the function's search space;
	\item \textit{Statistical Parameters about the solutions:} Mean and Median of the solutions are analysed to provide insights about the algorithms;
	\item Iterations to converge, function evaluations and function values.
\end{itemize}
\section{2D Function versions}
\label{functions2D}

@@

The minimal. maximal, mean and median values of the solutions are shown in Table~\ref{function_values:ackley}:

@@

The isolines of Ackley Function and convergence lines for all algorithms are shown in Figure~\ref{fig:ackley}:\begin{figure}[H]
\centering
\caption{Isolines and convergence line for Ackley Function}
\label{fig:ackley}
\includegraphics[scale=0.5]{images/ackley.jpg}
\end{figure}
The isolines for Ackley Function shows that there are at least 4 pitfalls near the optimal solution, so, the search space needs to be reduced to $[-0.05, 0.05]$ to avoid wrong convergence. All the algorithms converges quite smoothly following a straight line\subsection{Beale Function}
\label{beale2D}

@@

For implementation, it was decided to put constraints on search space, because for the original search space of some functions, like Ackley and Rastrigin, there are other minimas that will confuse every algorithm tested for this paper, and there is no sense on test a local minima procedure on these conditions. To restrain the initial solution to some region in the search space was a cheap and ready solution to avoid the algorithms to stop because of stumbling upon the wrong minimum.

@@

The \textit{Davidon–Fletcher–Powell} Algorithm is not implemented in \textit{SciPy} but its equations shows that it is the dual of BFGS, so, it was easy to implement using some \textit{SciPy} code for BFGS algorithm. In the DFP procedure, there is a Hessian inversion to be evaluated every iteration, then, numerical imprecision would be taken into consideration evaluating its results. While in some cases it has a superb performance (Like Ackley Function), it has very poor performance in Rosenbrock Function and even in Beale Function. The increase of dimensions makes DFP suffer of numerical imprecision due to the Hessian inversion, except for Rastrigin, and it is the loser algorithm in other functions.

@@

In fact, there are no Quasi-Newton algorithms that solver the minima problem for an any function. From the standpoint of performance and convergence, all the algorithms have an average performance and a reasonable convergence, although we have to restrain the search space to a small area near the optimal solution.

>>>>

\begin{itemize}
	\item \textit{Convergence: } the convergence of an algorithm is classified as \textit{good}, \textit{poor} and \textit{divergence}.
	\subitem \textit{Good convergence: } the final solution is really near the optimal solution within the max iterations;
	\subitem \textit{Poor convergence: } the final solution is far from the optimal solution for any reason;
	\subitem \textit{Divergence: }  the final solution is out of the function's search space;
	\item \textit{Statistical Parameters about the solutions:} Mean and Median of the solutions are analysed to provide insights about the algorithms;
	\item Iterations to converge, function evaluations and function values.
\end{itemize}

\section{2D Function versions}
\label{functions2D}

@@

The minimal, maximal, mean and median values of the solutions are shown in Table~\ref{function_values:ackley}:

@@

The isolines of Ackley Function and convergence lines for all algorithms are shown in Figure~\ref{fig:ackley}:\begin{figure}[H]
\centering
\caption{Isolines and convergence line for Ackley Function}
\label{fig:ackley}
\includegraphics[scale=0.5]{images/ackley.jpg}
\end{figure}
The isolines for Ackley Function shows that there are at least 4 pitfalls near the optimal solution, so, the search space needs to be reduced to $[-0.05, 0.05]$ to avoid wrong convergence. All the algorithms converges quite smoothly following a straight line

\subsection{Beale Function}
\label{beale2D}

@@

For implementation, it was decided to put constraints on search space, because for the original search space of some functions, like Ackley and Rastrigin, there are other minimas that will confuse every algorithm tested for this paper, and there is no sense on test a local minima procedure on these conditions. Restraining the initial solution to some region in the search space was a cheap and ready solution to avoid the algorithms to stop because of stumbling upon the wrong minimum.

@@

The \textit{Davidon–Fletcher–Powell} Algorithm is not implemented in \textit{SciPy} but its equations shows that it is the dual of BFGS, so, it was easy to implement using some \textit{SciPy} code for BFGS algorithm. In the DFP procedure, there is a Hessian inversion to be evaluated every iteration, then, numerical imprecision would be taken into consideration evaluating its results. While in some cases it has a superb performance (Like Ackley Function), it has very poor performance in Rosenbrock Function and even in Beale Function. The increase of dimensions makes DFP suffer of numerical imprecision due to the Hessian inversion, except for Rastrigin, and it is the loser algorithm in other test functions used for this work.

@@

In fact, there are no Quasi-Newton algorithms that solves the minima problem for an arbitrary function. From the standpoint of performance and convergence, all the algorithms have an average performance and a reasonable convergence, although we have to restrain the search space to a small area near the optimal solution.
